### 2025-02-06 paper/appendix.tex
Move section 3 to appendix

\onecolumn
\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge \\ (Supplementary Material)}
\maketitle
\appendix

\section{Prompts Used for LLM}
\label{section:llms}

\begin{figure}[ht!]
	\centering
	\begin{verbatim}
You are an expert in Social Science. Following are the descriptions of two variables:

<A>: {description of variable A}
<B>: {description of variable B}

Which of the following two options is the most likely causal direction between these 
variables:

1. <A> causes <B>
2. <B> causes <A>

Return a single letter answer between the choices above; Do not provide any reasoning 
in the answer; Do not add any text formatting to the answer.
	\end{verbatim}
	\caption{Prompt used for the LLM. Here the variable descriptions are replaced with description provided in Fig.~\ref{fig:var_description}}
	\label{fig:prompt}
\end{figure}

\begin{figure}[ht!]
	\begin{verbatim}
          Age: The age of a person
    Workclass: The workplace where the person is employed such as Private industry, 
     	       or self employed
    Education: The highest level of education the person has finished.
MaritalStatus: The marital status of the person
   Occupation: The kind of job the person does. For example, sales, craft repair, 
   		clerical.
 Relationship: The relationship status of the person.
         Race: The ethnicity of the person.
          Sex: The sex or gender of the person.
 HoursPerWeek: The number of hours per week the person works.
NativeCountry: The native country of the person.
       Income: The income i.e. amount of money the person makes.
	\end{verbatim}
	\caption{Variable descriptions used for prompting the LLM}
	\label{fig:var_description}
\end{figure}

\section{A Generalized Measure of Conditional Association}
\label{sec:mixed_association}

In the previous section, we discussed how different types of variables require
different measures of conditional association. However, there is no unified
measure that handles mixed data. In this section, we introduce a measure of conditional
association for mixed data by extending the concept of partial correlation
coefficient (commonly used for continuous variables) to mixed data. Similar to
partial correlation coefficient, our method integrates a mixed data
residualization method \citep{Ankan2023} with Pillai's Trace
\citep{Pillai1955}, a multivariate measure of association based on canonical
correlations.

Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
$, our goal is to estimate the conditional association $ \phi_{X, Y; \bm{Z}} $. 
In the first step, we compute the residuals $ R_X $ and $ R_Y $ for variables
$ X $ and $ Y $ respectively. Depending on the type of variable the residual
is computed as follows:

\begin{enumerate}
	\item \textbf{Continuous:} We train a model, $ E_X: x \sim
		\bm{z} $. The residuals are then computed by taking the difference
		between the true and the predicted values using $ E_X $. 
		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
	\item \textbf{Ordinal:} We start by training a probability estimator, $
		p_X: x \sim \bm{z} $, and then use the estimated probabilities, 
		$ \hat{p}_X(x) $ to compute the residuals:
		$$ R_{x_i} = \hat{p}_X(X < x_i) - \hat{p}_X(X > x_i) $$
	\item \textbf{Categorical:} We again start by training a probability
		estimator $ p_X: x \sim \bm{z} $, and obtain probability
		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
		categorical variable, resulting in a binary vector and then
		compute the residuals as follows: 
		$$ R_{x_i} = x_i - \hat{p}_X(\bm{z}_i) $$
\end{enumerate}

We have the option here to choose the estimators based on the characteristics
such as distribution, type of relationship, and so on of our dataset.
Non-parametric ensemble estimators such as Random Forest and XGBoost are robust 
for diverse data types and complex relationships. For linear relationships,
simpler models such as linear regression and its variants may suffice.

We repeat the above residualization step for both the variables $ X $ and $ Y $
obtaining residual matrices $ R_x $ and $ R_y $. The type of variable determines
the shape of these matrices. If the variable is continuous or ordinal, the 
residual matrix is of shape $ (n \times 1 ) $ and if the variable is categorical,
we get a residual matrix of shape $ (x \times (k - 1)) $, where $ k $ is the number
of categories of the variable.

The second step is to quantify the association between these residual matrices.
For this purpose we use canonical correlations \citep{Hotelling1936} that have been widely used to
measure the association between sets of random variables.

\begin{definition}
	Given two sets of random variables $ \bm{U} = (U_1, U_2, \cdots, U_p) $
	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
	them, $\rho_{\bm{U}, \bm{V}} $ is defined as:

	\begin{equation}
		% \nu_{\bm{U}, \bm{V}}= \max_{a, b} \mathrm{corr}(a^T \bm{U}, b^T \bm{V})
		\rho_{\bm{U}, \bm{V}} = \max_{a, b} \frac{a^T \Sigma_{\bm{UV}} b}{\sqrt{a^T \Sigma_{\bm{UU}} a \cdot b^T \Sigma_{\bm{VV}} b}}
	\end{equation}

	where $ a $ and $ b $ are vectors of coefficients that maximize the correlations
	between the linear combinations of $ a^T \bm{U} $ and $ b^T \bm{V} $.
\end{definition}

Canonical correlations generalize the concept of correlation coefficients to
multi-dimensional variables. It finds orthogonal linear transformations $ a $
and $ b $ that maximized the correlation between the transformed variables $
a^T \bm{U} $ and $ b^T \bm{V} $. This yields a vector of correlation
coefficient values of size $ \min(p, q) $, representing the correlation
coefficient of each pair of transformed variables. Notably, Pearson's
correlation coefficient is a special case of canonical correlations when $ p =
q = 1 $.

Several measures of association have been derived from canonical correlations, such as:
\begin{itemize}
	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \rho_i^2) $
	\item Roy's Largest Root: $ \theta = \max_i(\rho_i^2) $
	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \rho_i^2 $
\end{itemize}

We use a normalized version of Pillai's Trace for our purpose, given as:

\begin{equation}
	\tau_{X, Y; \bm{Z}} = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\rho_{R_x, R_y})_i^2
\end{equation}

We use Pillai's Trace because: 1) It uses all the canonical correlation values,
capturing the full extent of association, 2) Its interpretation is similar to
Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
$ signifies perfect linear relationship. We use a normalized version because
when comparing between pair of variables with different number of categories.
Our proposed measure has several desirable properties that make it well-suited
for our application:

\begin{enumerate}
	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
		association) to $ 1 $ (perfect linear association), simplifying
		interpretation.
	% \item \textbf{Independent of Sample Size: } 
	\item \textbf{Invariant to the Dummy Encoding: } For categorical
		variables, residuals are computed using dummy encoding. Since
		canonical correlations identifies linear combinations across
		columns, this measure is invariant to the specific dummy
		encoding scheme used.
	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
		When both $ X $ and $ Y $ are continuous variables, our measure
		is equivalent to the absolute value of partial correlation coefficient.
	\item \textbf{Equivalent to polychoric and polyserial correlation for
			ordinal and continuous variables: }
		Under the assumption that the ordinal variable is generated by
		discretizing an underlying gaussian continuous variable, this
		effect size is equivalent to polychoric and polyserial
		correlation. As both of them recover the Pearson correlation
		coefficient.
		\todo[inline]{Verify if this is correct}
	\item \textbf{Related to Cram\'er's V: } 
\end{enumerate}

In essence, our measure of conditional association extends existing metrics to mixed data, providing 
a unified measure that is interpretable.